A deeper look into bug fixes: Patterns, replacements, deletions, and additions

### 2016-02-19 MSRChallenge2016.tex
four pages, legible figures

\maketitle
\begin{abstract}
  Many implementations of research techniques that automatically repair software
  bugs target programs written in C. Work that targets Java often begins from, or at
  least compares to, direct translations of such techniques to a Java context.
  However, Java and C are very different languages, and these differences should be
  studied and understood if existing approaches are to be profitably translated
  between them.  We conducted a large-scale study of bug-fixing commits
  to Java projects, focusing on assumptions underlying common
  search-based repair approaches.  We make observations that can be
  leveraged to guide high quality automatic software repair to target Java specifically,
  including common and uncommon statement replacements as conducted by human
  developers and the applicability of previously-proposed patch construction
  operators in the Java context. 
\end{abstract}

@@

There has been considerable attention paid to the problem of automatic
program repair in recent
years~\cite{kim2013,legoues2012,pan2009,
  Mechtaev15,Long2016}, among many others),
%
One broad class of techniques in this space take a generate-and-validate
approach, which involves generating a large number of candidate patches using a
pre-defined set of mutation operators, and then validating correctness with
respect to a set of test cases.  Mutation operators range from simple, coarse
granularity statement-level mutations to human-constructed templates learned
from a large corpus of previous human bug-fixing commits.

A considerable proportion of this works targets programs written in C.
Researchers targeting Java often start from or compare against imperfect
baselines, that is, Java-based implementations of techniques originally
implemented for C~\cite{nopol,kim2013}. % Even PAR~\cite{kim2013}, perhaps the
% best-known previously-proposed Java-based program repair technique, both starts
% from and is directly compared to a Java implementation of GenProg~\cite{genprog}, which
% targets the C programming language.  
%
This is an oversight because Java and C are very different languages.
Consider GenProg~\cite{legoues2012}, which
combines statement-level mutations into patches to address a particular
defect. Which ``Statements'' (a semantic unit in C) should be manipulated in Java?
Can we learn
intelligent rules to inform candidate modifications
(cf. Prophet~\cite{Long2016}, for C)?  What types of semantic scoping
should limit mutations, given the relative stringence of the Java
compiler?

In this paper, we study bug-fixing commits to Java programs, taking as a subject
of study several million human-made bug fixes from Github. We study broad
characteristics of these changes; the applicability of previously-proposed
Java repair templates~\cite{kim2013}; and the nature
of additions and replacements of Java statements in bug-fixing commits. We make
observations to directly guide future research in automatic repair of Java
programs, to increase the success rate of such techniques and the degree to
which the patches are human-like and therefore more readable and maintainable by
human developers. This is important, because patch quality is an important
concern in this research area~\cite{Qi15}.

@@

\noindent\textbf{Related work.} Zhong and Su~\cite{zhong2015} study
a similar set of questions in C.  Beyond focusing on Java, we focus more on
replacements, ommitted from that previous analysis, and target a
larger dataset~\cite{dyer2013}.  Barr et al.~\cite{Barr14fse} studied changes to
Java programs to understand the ``Plastic Surgery Hypothesis'' underlying
certain types of program repair, an important but orthogonal
concern. Kim et al.~\cite{kim2013} manually analyze 
changes to Java programs to inform an automated repair
technique and show that doing so results in higher-quality repairs; 
we attempt to validate their templates heuristically on
a different dataset, and take their results as
motivation for understanding human repairs (as it may result in better patches).
Long and Rinard learn probabilistic models from bug fixes to C
code~\cite{Long2016}. Our analysis is significantly less precise and does not
inform a new repair technique.  Rather, it serves as a starting
point for research investigating the repair of bugs in Java code.

@@

We begin by broadly characterizing the bug-fixing changes to Java code in the
dataset.  We perform our analysis on the \emph{2015 September/Github} dataset on
Boa, including $7,830,023$ Java projects with $23,229,406$ revisions. Boa
identifies $4,590,679$ as bug fixing.

\vspace{1ex}
\noindent\textbf{How many files are changed to fix a bug?}
%
We consider a file changed if it is new, modified, or
deleted in a commit. In total, $52,052,571$ files were changed. Thus, on average each bug
fixing revision changes $11.3$ files.  This is surprisingly high:
most automatic program repair techniques assume that bugs are local.
We hypothesize several explanations. Other
software artifacts (e.g.  documentation) may be updated after bug  
fixing, or fixing commits may contain changes
unrelated to fixing bug (e.g. refactoring, adding new features by
creating new files). To better understand this
phenomenon, we ask:

\vspace{1ex}
\noindent\textbf{What are the file types of those changed files?}
%
To answer this question, we iterate over all changed files in each bug fixing
revision and count the total number for each file type. Results are shown in
Table~\ref{tbl:fileType}.

\begin{table}
\centering
  \begin{tabular}{ l | r  r }
  \toprule
  File Kind & Total & Average \\ 
  \midrule
  Binary & 752,945 & 0.16 \\ 
  Java (ERROR) & 2,073,558 & 0.45 \\ 
  Java (JLS2) & 2,607,413 & 0.57 \\ 
  Java (JLS3) & 15,748,967 & 3.43 \\
  Java (JLS4)  & 83,798 & 0.02 \\ 
  Text & 541,023 & 0.12 \\ 
  XML & 6,818,299 & 1.49 \\ 
  UNKNOWN & 23,426,568 & 5.10 \\ 
\bottomrule
  \end{tabular}
  \caption{File Types of Changed Files}
  \label{tbl:fileType}
\end{table}

JAVA (ERROR) files are those that Boa failed to parse. JLS2,
JLS3, and JLS4 represent Java files written in different J2SE versions. From
Table~\ref{tbl:fileType}, we see that Text and Binary files are
changed least frequently. This is unsurprising, since text are often
documentation files, and binaries should only be changed rarely. If we consider all 4 kinds of Java
files as one kind, on average, each bug fixing revision changes $4.47$ Java
files. This still does not conform to our assumption of bug locality. XML
files in Java projects usually represents build files; changing a build
file is unsurprising.  Rather more surprising is how frequently UNKNOWN files
are changed. Given that all the projects are Java projects, 
it is unclear whether these are written in other proramming
languages. We hypothesize they might be test resources.
To better understand what changes are usually made to Java source code
files, we next ask:

\vspace{1ex}
\noindent\textbf{How often do developers introduce new classes, methods, fields,
	local variables in
bug fixing commits?} 
%
This question is interesting because most automatic program repair approaches
target C, and do not consider
object-oriented concerns, like fields or classes.

@@

\begin{table}
\centering
  \begin{tabular}{l | r  r }
  \toprule
  Introducing & Total Number & Average \\ \midrule
  Class & 729,201 & 0.16 \\ 
  Methods & 3,186,867 & 0.69 \\ 
  Fields & 6,076,646 & 1.32 \\ 
  Variables & 924,259 & 0.2 \\ 
\bottomrule
  \end{tabular}
  \caption{Introducing new classes, methods, fields, local variables to fix bugs}
  \label{tbl:new}
\end{table}

@@

One of the best-known program repair techniques to target Java is
PAR~\cite{kim2013}.  Par searches for specific patterns
and modifies java code according to predefined templates, constructed by humans
to cover a large set of bug fixes from an existing corpus.  Par serves as an
important source of possible 
mutations to study to inform future directions. 
We therefore search for and analyze most of the bug fixing patterns proposed by
Kim et al. and estimate their prevalence in this
dataset.

@@

\subsection{Detecting PAR patterns in Boa}

Boa's capabilities are powerful, but limited in the precision it enables in
our detection of Par bug-fixing patterns. For example, it is unable to directly perform a
diff between two file versions. Thus, rather than finding the exact count of
bug fixing patterns, we find an approximate count by processing pre- and
post- fix files separately.  Additionally, there are
two of the originally proposed 10 patterns that cannot be easily detected by
Boa, as we describe below. 
%
The patterns we search for are as follows, including our mechanism for
detecting them using Boa:

\vspace{1ex}
\noindent {\bf Altering method parameters (AMP)}.  This pattern changes 
input method parameters. %(\texttt{obj.method(v1,v2) $\rightarrow$
%obj.method(v1,v3)}). 
To detect this pattern, for both pre-fix and post-fix version
of a buggy file, we create a custom method call signature that
contains the name, literal parameters, and variable parameters (more complex
expressions are listed as OTHER).
%  This is due to inexistence of
% function that can print Abstract Syntax Tree (AST) back to source code in
% Boa. 
We discard method signatures that appear both in pre-fix and post-fix
versions, and then identify methods with the same name and the
same number of parameters, but different parameter signatures in pre- and
post-fix versions.% If it exists, we consider that AMP pattern is found.

\vspace{1ex}
\noindent {\bf Calling another method with the same parameters (MSM)}.  This
pattern changes a method name. % (\texttt{obj.method1(param) $\rightarrow$
%obj.method2(param)}).  
To detect this pattern, we create method signature similar
to the above for the AMP pattern, looking for methods
with the exact same parameter signature but different names in pre- and
post-fix versions. %If it exists, we consider that MSM pattern is found.

\vspace{1ex}
\noindent
{\bf Calling another overloaded method with one more parameter (COM)}. This
pattern adds one more parameter to a method. %(\texttt{obj.method(v1) $\rightarrow$ obj.method(v1,v2)}).
We create method signature similar to the above, seeking methods with
  the same name, one more parameter in the post-fix version.

\vspace{1ex}
\noindent
 {\bf Changing a branch condition (CBC)}. This pattern adds or removes a condition.
%(e.g., \texttt{if(a != b) $\rightarrow$ if(a != b \&\& c == 0)}).
  To detect CBC pattern, we count the number of { \em logical and} and {\em
    logical or} inside if conditional expression sfor both pre-fix and post-fix
  version. We assume that the addition/removal of logical operators indicate
  addition/removal of condition. If there is a larger count in post-fix version,
  we consider the CBC pattern is found.

\vspace{1ex}
\noindent
{\bf Initializing an object (IAO)}. This pattern adds a initialization to object
declaration.  We count the number of NEW expressions in variable declarations
for both pre- and post-fix versions. If there is a larger count in post-fix
version, we consider that IAO pattern is found.

\vspace{1ex}
\noindent
 {\bf Adding a null checker (ANC)}.  This pattern inserts a condition to check whether an
    object is null.
%	{\bf Example:} obj.m1() $\rightarrow$ if(obj!=null){obj.m1()}
  To detect ANC pattern, we count the number of if conditional expressions that
  contains {\em !=null} or {\em ==null} for both pre-fix and post-fix
  version. If there is a larger count in post-fix version,
  we consider that ANC pattern found.

@@

Table~\ref{tab:freqpattern} estimates the prevelance of PAR templates in the
Java dataset. The most common pattern we observed is CBC; and the least common
pattern is AOB. If we assume that these patterns never appear together, they
cover 14.78\% of buggy files. Kim et al. mentioned that the patterns that they
use cover almost 30\% of real patches~\cite{kim2013}. Although we only detect 8
out of 10 bug fixing patterns, it is unlikely the final two will cover the
difference. Moreover, our detection is imprecise, and likely represents an upper
bound.  This suggests that these fixing patterns might not generalize
beyond an initial dataset, and that more work remains to categorize the programmer's fixing behavior.

\section{Statement-level mutations}
\begin{table*}
	\centering
	\resizebox{15cm}{!}{
	%\renewcommand{\arraystretch}{0.9}% Tighter
		\begin{tabular}{|l|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|} \hline
			&Assert&Break&Continue&Do& For&If&Label&Return&Case&Switch& Synch&Throw&Try&TypeDecl&While
			\\ \hline

\texttt{Assert}	&-&7.48&3.76&0.53&8.30&23.05&0.31&20.04&4.90&4.62&1.30&13.50&7.23&0.03&4.95
\\ \hline
\texttt{Break} &1.00&-&4.08&0.60&9.93&26.03&0.13&25.39&2.48&1.57&1.79&8.39&11.73&0.10&6.77
\\ \hline
\texttt{Continue} &1.74&9.42&-&1.28&11.39&18.25&0.35&22.60&3.80&2.85&2.17&8.98&9.42&0.11&7.63
\\   \hline
\texttt{Do} &0.81&5.26&6.60&0.00&9.44&14.21&0.18&15.86&3.73&1.67&1.97&5.88&6.39&0.03&27.98
\\ \hline
\texttt{For} &0.86&6.28&3.19&0.79&-&22.89&0.09&21.08&5.01&3.34&1.87&10.01&10.71&0.08&13.79
\\ \hline
\texttt{If}&1.64&8.43&2.87&0.60&13.49&-&0.24&26.46&7.45&4.80&2.85&9.89&15.11&0.08&6.11
\\ \hline
\texttt{Label} &1.30&8.33&7.86&1.11&5.18&22.85&-&15.17&3.05&2.04&14.62&10.45&4.16&0.09&3.79
\\ \hline
\texttt{Return} &1.13&9.41&3.11&0.49&13.33&27.24&0.24&-&5.59&3.65&2.55&14.91&12.61&0.12&5.61
\\ \hline
\texttt{Case} &0.78&2.84&2.84&0.39&10.27&31.79&0.16&22.40&-&0.46&2.07&7.37&11.69&0.08&6.87
\\ \hline
\texttt{Switch}&1.14&2.72&3.80&0.55&11.07&34.14&0.13&21.86&0.75&-&1.53&8.65&9.02&0.05&4.58
\\ \hline
\texttt{Synch} &0.80&6.57&2.28&0.43&10.21&24.18&0.05&19.77&6.35&2.07&-&9.16&12.16&0.04&5.93
\\ \hline
\texttt{Throw} &2.11&6.57&2.58&0.48&11.87&18.84&0.17&32.28&4.64&3.30&2.74&-&10.08&0.07&4.27
\\ \hline
\texttt{Try}&0.71&7.41&3.02&0.66&11.73&27.75&0.11&23.24&5.63&2.65&2.58&8.99&-&0.09&5.42
\\ \hline
\texttt{TypeDecl} &0.00&4.51&7.52&1.00&10.28&21.05&0.50&17.79&6.02&1.75&2.01&9.27&11.53&-&6.77
\\ \hline
\texttt{While}& 0.72&8.02&3.82&1.96&23.16&19.78&0.12&16.48&6.56&3.09&1.64&6.81&7.80&0.04&-
			\\ \hline
		\end{tabular}
		}
		\caption{Likeliness to replace the statement in the first column for the statement in the top row, in Java}\label{tab:likeliness}
\end{table*}

\begin{figure*}[!t] \centering \includegraphics[height=0.23\textheight,
	width=0.8\textwidth]{StmtType} \caption{For each kind of statement, how many
	bug fixing revisions add/delete statements of the same kind?}
	\label{fig:StmtType} \end{figure*}

>>>>

\maketitle
\begin{abstract}
  Many implementations of research techniques that automatically repair software
  bugs target programs written in C. Work that targets Java often begins from, or at
  least compares to, direct translations of such techniques to a Java context.
  However, Java and C are very different languages, and so Java should be
  studied to inform understanding of how to translate existing approaches to target it.
  We conduct a large-scale study of bug-fixing commits
  to Java projects, focusing on assumptions underlying common
  search-based repair approaches.  We make observations that can be
  leveraged to guide high quality automatic software repair to target Java specifically,
  including common and uncommon statement replacements as conducted by human
  developers and the applicability of previously-proposed patch construction
  operators in the Java context. 
\end{abstract}

@@

There has been considerable attention paid to automatic
program repair in recent
years~\cite{kim2013,legoues2012,pan2009,
  Mechtaev15,Long2016}, among many others),
%
One broad class of techniques in this space take a generate-and-validate
approach: \emph{generating} a large number of candidate patches using a
pre-defined set of mutation operators, and then \emph{validating} correctness with
respect to a set of test cases.  Mutation operators range from simple, coarse
granularity statement-level mutations to human-constructed templates learned
from a large corpus of previous human bug-fixing commits.

A considerable proportion of this works targets programs written in C.
Indeed, researchers targeting Java often start from or compare against
Java-based implementations of techniques originally
implemented for C~\cite{nopol,kim2013}. % Even PAR~\cite{kim2013}, perhaps the
% best-known previously-proposed Java-based program repair technique, both starts
% from and is directly compared to a Java implementation of GenProg~\cite{genprog}, which
% targets the C programming language.  
%
This is an oversight because Java and C are very different languages.
Consider GenProg~\cite{legoues2012}, which
combines statement-level mutations into patches to address a particular
defect. Which ``Statements'' (a semantic unit in C) should be manipulated in Java?
Can we learn
intelligent rules to inform candidate modifications
(cf. Prophet~\cite{Long2016}, for C)?  What types of semantic scoping
should limit mutations, given the relative stringence of the Java
compiler?

In this paper, we study bug-fixing commits to Java programs, 
taken from several million human-made bug fixes from Github. We study broad
characteristics of these changes; the applicability of previously-proposed
Java repair templates~\cite{kim2013}; and the nature
of additions and replacements of Java statements in bug-fixing commits. We make
observations to directly guide future research in automatic repair of Java
programs, to increase the success rate of such techniques and the degree to
which the patches are human-like and therefore more readable and maintainable by
human developers. This is important, because patch quality is an important
concern in this area~\cite{Qi15}.

@@

\vspace{1ex}
\noindent\textbf{Related work.} Zhong and Su~\cite{zhong2015} study
a similar set of questions as we do, for C.  Beyond focusing on Java in a larger dataset~\cite{dyer2013}, we also study
replacements, ommitted from that analysis.  Barr et al.~\cite{Barr14fse} studied changes to
Java programs to understand the ``Plastic Surgery Hypothesis'' underlying
certain types of program repair, an orthogonal
concern. Kim et al.~\cite{kim2013} manually analyze 
changes to Java programs to inform an automated repair
technique, and show that doing so results in higher-quality repairs; 
we study their templates heuristically on
a different dataset. Their results motivate
studies of human repairs, as it may result in better patches.
Long and Rinard learn probabilistic models from bug fixes to C
code~\cite{Long2016}. Our analysis is significantly less precise and does not
inform a new technique.  Rather, it serves as a starting
point for research on the repair of Java bugs.

@@

First, we broadly characterize bug-fixing commits to Java code.  In this paper,
we study the \emph{2015 September/Github} dataset on Boa, including $7,830,023$
Java projects with $23,229,406$ revisions. Boa identifies $4,590,679$ as bug
fixing.

\vspace{1ex}
\noindent\textbf{How many files are changed to fix a bug?}
%
We consider a file changed if it is new, modified, or
deleted in a commit. In total, $52,052,571$ files were changed, an average of 
$11.3$ files per bug-fixing commit.  This is surprisingly high:
most automatic program repair techniques assume that bugs are local.
We hypothesize several explanations: Other
software artifacts (e.g.  documentation) may be updated after bug  
fixing, or fixing commits may contain changes
unrelated to fixing the bug (e.g. refactoring, feature addition). 
To better understand this
phenomenon, we ask:

\begin{table}
\centering
  \begin{tabular}{ l | r  r }
  \toprule
  File Kind & Total & Average \\ 
  \midrule
  Binary & 752,945 & 0.16 \\ 
  Java (ERROR) & 2,073,558 & 0.45 \\ 
  Java (JLS2) & 2,607,413 & 0.57 \\ 
  Java (JLS3) & 15,748,967 & 3.43 \\
  Java (JLS4)  & 83,798 & 0.02 \\ 
  Text & 541,023 & 0.12 \\ 
  XML & 6,818,299 & 1.49 \\ 
  UNKNOWN & 23,426,568 & 5.10 \\ 
\bottomrule
  \end{tabular}
  \caption{File Types of Changed Files}
  \label{tbl:fileType}
\end{table}
\vspace{1ex}
\noindent\textbf{What are the types of those changed files?}
%
To answer this question, we iterate over all changed files in each bug fixing
revision and count the total number for each file type. Results are shown in
Table~\ref{tbl:fileType}.
%
JAVA (ERROR) files are those that Boa failed to parse. JLS2,
JLS3, and JLS4 represent Java files written in different J2SE versions.  We see that text and Binary files are
changed least frequently. This is unsurprising, since text files are often
documentation, and binaries should only be changed rarely. If we consider all 4 kinds of Java
files as one kind, on average, each bug fixing revision changes $4.47$ Java
files. This still does not conform to our assumption of bug locality. XML
files in Java projects usually represent build files; changing a build
file is unsurprising.  Rather more surprising is how frequently UNKNOWN files
are changed. Given that all the projects are Java projects, 
it is unclear whether these are written in other proramming
languages. We hypothesize they might be test resources.
\todo{possibly mention median, if Chu Pan gets that information}
To better understand what changes are usually made to Java source code
files, we next ask:

\vspace{1ex}
\noindent\textbf{How often do developers introduce new classes, methods, fields,
	local variables in
bug fixing commits?} 
%
This question is interesting because most automatic program repair approaches
do not create methods or variable declarations, nor do most explicitly consider
object-oriented features, like fields or classes.

@@

\begin{table}
\centering
  \begin{tabular}{l | r  r }
  \toprule
  Introducing & Total Number & Average \\ \midrule
  Class & 729,201 & 0.16 \\ 
  Methods & 3,186,867 & 0.69 \\ 
  Fields & 6,076,646 & 1.32 \\ 
  Variables & 924,259 & 0.2 \\ 
\bottomrule
  \end{tabular}
  \caption{Introducing new classes, methods, fields, local variables to fix bugs}
  \label{tbl:new}
\vspace{-0.5cm}
\end{table}

@@

One of the best-known program repair techniques to target Java is
PAR~\cite{kim2013}.  PAR
modifies Java code according to predefined templates, constructed by humans
to cover a large set of bug fixes from an existing corpus.  These templates provide
important source of possible 
mutations to study to inform future directions. 
We therefore search for most of the PAR bug fixing patterns,
estimating their prevalence in this
dataset.

@@

\subsection{Detecting PAR templates using Boa}

Boa's capabilities are powerful, but limited in the precision it enables in
our detection of Par bug-fixing patterns. For example, it cannot directly \texttt{diff}
two files versions. Thus, rather than finding the exact count of
bug fixing patterns, we approximate by processing pre- and
post- fix files separately.  Additionally, there are
two of the 10 patterns that cannot be easily detected by
Boa, as we describe below. 
%
The patterns we search for are as follows, including our mechanism for
detecting them:

\vspace{1ex}
\noindent {\bf Altering method parameters (AMP)}.  This template changes 
input method parameters. %(\texttt{obj.method(v1,v2) $\rightarrow$
%obj.method(v1,v3)}). 
To detect this pattern, for both pre-fix and post-fix versions
of a buggy file, we create a custom method call signature that
contains the name, literal parameters, and variable parameters (complex
expressions are listed as OTHER).
%  This is due to inexistence of
% function that can print Abstract Syntax Tree (AST) back to source code in
% Boa. 
We discard method signatures that appear both in pre-fix and post-fix
versions, and then identify methods with the same name and the
same number of parameters, but different signatures between versions.

\vspace{1ex}
\noindent {\bf Calling another method with the same parameters (MSM)}.  This
template changes a method name. % (\texttt{obj.method1(param) $\rightarrow$
%obj.method2(param)}).  
To detect it, we create a method signature similar
to the above for AMP, looking for methods
with the exact same parameter signature but different names in pre- and
post-fix versions. %If it exists, we consider that MSM pattern is found.

\vspace{1ex}
\noindent
{\bf Calling another overloaded method with one more parameter (COM)}. This
template adds a parameter to a method. %(\texttt{obj.method(v1) $\rightarrow$ obj.method(v1,v2)}).
We create a method signature similar to the above, seeking methods with
  the same name, one more parameter in the post-fix version.

\vspace{1ex}
\noindent
 {\bf Changing a branch condition (CBC)}. This template adds or removes a condition.
%(e.g., \texttt{if(a != b) $\rightarrow$ if(a != b \&\& c == 0)}).
We count the number of { \em logical and} and {\em
    logical or} inside if conditional expression sfor both pre-fix and post-fix
  version. We assume that the addition/removal of logical operators indicate
  addition/removal of condition. If there is a larger count in post-fix version,
  we consider the CBC pattern is found.

\vspace{1ex}
\noindent
{\bf Initializing an object (IAO)}. This template adds a initialization to object
declaration.  We count the number of NEW expressions in variable declarations
for both pre- and post-fix versions. If there is a larger count in post-fix
version, we consider that IAO pattern is found.

\vspace{1ex}
\noindent
 {\bf Adding a null checker (ANC)}.  This template inserts a condition to check whether an
    object is null.
%	{\bf Example:} obj.m1() $\rightarrow$ if(obj!=null){obj.m1()}
  To detect ANC pattern, we count the number of if conditional expressions that
  contains {\em !=null} or {\em ==null} for both pre-fix and post-fix
  version. If there is a larger count in post-fix version,
  we consider that ANC pattern found.

@@

Table~\ref{tab:freqpattern} estimates the prevelance of PAR templates in the
Java dataset. The most common pattern we observed is CBC; and the least common
pattern is AOB. If we assume that these patterns never appear together, they
cover 14.78\% of buggy files. Kim et al. mentioned that the patterns that they
use cover almost 30\% of real patches~\cite{kim2013}. Although we only detect 8
out of 10 bug fixing patterns, it is unlikely the final two will cover the
difference. Moreover, our detection is imprecise, and likely represents an upper
bound.  This suggests that these fixing patterns might not generalize
beyond an initial dataset, and that more work remains to expressively characterize human bug-fixing behavior.

\section{Statement-level mutations}
\begin{table*}
	\centering
	\resizebox{\textwidth}{!}{
	%\renewcommand{\arraystretch}{0.9}% Tighter
		\begin{tabular}{l|rrrrrrrrrrrrrrr} \toprule
			&\texttt{Assert}&\texttt{Break}&\texttt{Continue}&\texttt{Do}&\texttt{For}&\texttt{If}&\texttt{Label}&\texttt{Return}&\texttt{Case}&\texttt{Switch}& \texttt{Synch}&\texttt{Throw}&\texttt{Try}&\texttt{TypeDecl}&\texttt{While}
			\\ \midrule

\texttt{Assert}	&-&7.48&3.76&0.53&8.30&23.05&0.31&20.04&4.90&4.62&1.30&13.50&7.23&0.03&4.95
\\
\texttt{Break} &1.00&-&4.08&0.60&9.93&26.03&0.13&25.39&2.48&1.57&1.79&8.39&11.73&0.10&6.77
\\
\texttt{Continue} &1.74&9.42&-&1.28&11.39&18.25&0.35&22.60&3.80&2.85&2.17&8.98&9.42&0.11&7.63
\\
\texttt{Do} &0.81&5.26&6.60&0.00&9.44&14.21&0.18&15.86&3.73&1.67&1.97&5.88&6.39&0.03&27.98
\\
\texttt{For} &0.86&6.28&3.19&0.79&-&22.89&0.09&21.08&5.01&3.34&1.87&10.01&10.71&0.08&13.79
\\
\texttt{If}&1.64&8.43&2.87&0.60&13.49&-&0.24&26.46&7.45&4.80&2.85&9.89&15.11&0.08&6.11
\\
\texttt{Label} &1.30&8.33&7.86&1.11&5.18&22.85&-&15.17&3.05&2.04&14.62&10.45&4.16&0.09&3.79
\\
\texttt{Return} &1.13&9.41&3.11&0.49&13.33&27.24&0.24&-&5.59&3.65&2.55&14.91&12.61&0.12&5.61
\\
\texttt{Case} &0.78&2.84&2.84&0.39&10.27&31.79&0.16&22.40&-&0.46&2.07&7.37&11.69&0.08&6.87
\\
\texttt{Switch}&1.14&2.72&3.80&0.55&11.07&34.14&0.13&21.86&0.75&-&1.53&8.65&9.02&0.05&4.58
\\
\texttt{Synch} &0.80&6.57&2.28&0.43&10.21&24.18&0.05&19.77&6.35&2.07&-&9.16&12.16&0.04&5.93
\\
\texttt{Throw} &2.11&6.57&2.58&0.48&11.87&18.84&0.17&32.28&4.64&3.30&2.74&-&10.08&0.07&4.27
\\
\texttt{Try}&0.71&7.41&3.02&0.66&11.73&27.75&0.11&23.24&5.63&2.65&2.58&8.99&-&0.09&5.42
\\
\texttt{TypeDecl} &0.00&4.51&7.52&1.00&10.28&21.05&0.50&17.79&6.02&1.75&2.01&9.27&11.53&-&6.77
\\
\texttt{While}& 0.72&8.02&3.82&1.96&23.16&19.78&0.12&16.48&6.56&3.09&1.64&6.81&7.80&0.04&-
			\\ \bottomrule
		\end{tabular}
		}
		\caption{Likelihood of replacing a statement type (row) by a statement of
          another type (column), for Java.}\label{tab:likeliness}
\end{table*}

\begin{figure*}[!t] \centering \includegraphics[width=\textwidth]{StmtType} \caption{For each kind of statement, how many
	bug fixing revisions add/delete statements of the same kind?}
	\label{fig:StmtType} \end{figure*}
